Automatically Splitting a Staged Lambda Calculus

### 2015-07-10 paper/relatedwork.tex

In the partial evaluation community, stage splitting has been investigated 
mostly as a preprocessing step for self-applied partial evaluators. 
For instance,~\cite{Mogensen89a} and~\cite{DeNiel91} defined splitting algorithms
for programs consistening of a set of first-order recursive functions.
Those works use a grammar-based binding-time specification that's 
relatively complicated compared to the type system of \lang.
This complication mostly arises from the decision to start with an untyped language:
one must first find a way to say ``argument $x$ will be a tuple'' before one can say
``the first component of $x$ is available at the second stage,'' 
and so much of their binding-time specification just amounts to a type system.
Formulations more like \lang\ (namely the two-level system of~\cite{NN92-twolevel}) were known at the time,
but~\cite{Mogensen89a} found these unsatisfactory because they could not describe data structures with mixed stage.
Davies essentially solved this limitation in~\cite{davies96} (pending our further tweak with $\curr$),
and the question of stage splitting as we have presented it has been ripe ever since.
Binding time analysis for a static higher-order language was visited in \cite{Mogensen89b},
but the implications for splitting were not explicitly considered.

>>>>

In the partial evaluation community, stage splitting has been investigated 
mostly as a preprocessing step for self-applied partial evaluators. 
For instance,~\cite{Mogensen89a} and~\cite{DeNiel91} defined splitting algorithms
for programs consisting of a set of first-order recursive functions.
Those works use a grammar-based binding-time specification that's 
relatively complicated compared to the type system of \lang.
This complication mostly arises from the decision to start with an untyped language:
one must first find a way to say ``argument $x$ will be a tuple'' before one can say
``the first component of $x$ is available at the second stage,'' 
and so much of their binding-time specification just amounts to a type system.
Formulations more like \lang\ (namely the two-level system of~\cite{NN92-twolevel}) were known at the time,
but~\cite{Mogensen89a} found these unsatisfactory because they could not describe data structures with mixed stage.
Davies essentially solved this limitation in~\cite{davies96} (pending our further tweak with $\curr$),
and the question of stage splitting as we have presented it has been ripe ever since.
Binding time analysis for a static higher-order language was visited in \cite{Mogensen89b},
but the implications for splitting were not explicitly considered.
